Internal models of type theory

### 2013-10-24 article-models-defn.tex
introducing strict compatibility properties for substitution and weakening
putting the new macros \ctxext, \ctxwk and \subst into use

\subsection{Extension}
We introduce extension which not only extends a context $\Gamma$ and a type
$A$ over it to a context $\ctxext{\Gamma}A$, but which also extends a type $A$
in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}P$ in context
$\Gamma$. We do this to ensure that all of type theory can be done in a context.
For instance, we could say (1) that a context in context $\Gamma$ is the same thing
as a type in context $\Gamma$; (2) When $A$ is a context in this sense, a type in
context $A$ is the same thing as a family $P$ over $A$ and (3) when $P$ is a type
in context $A$ in this sense, a term of $P$ keeps its original meaning.

Note that by introducing extension on the level of types and families,
we introduce $\Sigma$-types at a very early stage.
However, we need substitution to make this precise.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}A}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}A}{\ctxext{\Delta}B}}\\
\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\ctxext{A}P}}
& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}A}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}P}{\ctxext{B}Q}}
\end{infarray}

\subsection{Weakening}
We first define weakening by a context $\Gamma$. Since weakening by $\Gamma$
should in principle be a functor, it acts on contexts, types and terms alike.
Note that it is because of the weakening $\ctxwk{\Gamma}{\Delta}$ of a context
$\Delta$ by $\Gamma$ that we can speak of context morphisms from $\Gamma$ to $\Delta$: they are the terms
of $\ctxwk{\Gamma}{\Delta}$.
\begin{infarray}{cc}
\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
\end{infarray}

A weakening operation is also defined for types. When $A$ and $B$ are both types
 in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$
 is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
 are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}y$ are the
 constant maps at $y$ from $A$ to $B$, for $y:B$.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}A}{\ctxwk{A}{B}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}B}{Q}}
{\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}B}{Q}{Q'}}
{\jtypeeq{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
\end{infarray}

\subsection{Substitution}
\begin{infarray}{cc}
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}A}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext({\Gamma}{A})P}{Q}}{\jtype{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext({\Gamma}{A})P}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}A}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{\Gamma}A}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext({\Gamma}{A}){P}}{Q}{g}}{\jterm{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{g}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext({\Gamma}{A})P}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
\end{infarray}

@@

Extension is compatible with extension:
\begin{infarray}{c}
\inference
{ \jtype{\ctxext{\Gamma}A}{P}
}{\jterm
  {\ctxext{\Gamma}({A}P)}
  {\ctxwk{\ctxext{\Gamma}({A}P)}{(\ctxext({\Gamma}{A})P)}}{\typefont{extext}_0(A,P)}
}\\
\inference
{ \jtype{\ctxext({\Gamma}{A})P}{Q}
}{\jterm
  {\ctxext{\Gamma}({A}({P}Q))}
  {\ctxwk{\ctxext{A}({P}Q)}{(\ctxext({A}{P})Q)}}
  {\typefont{extext}_1(A,P,Q)}
}
\end{infarray}

Weakening is compatible with extension:
\begin{infarray}{c}
\inference{
  \jctx{\Gamma}
  \qquad
  \jtype{\Delta}{B}}
  {\jterm
    {\ctxext{\Gamma}\ctxwk{\Gamma}{(\ctxext{\Delta}B)}}
    {\ctxwk{\ctxwk{\Gamma}{(\ctxext{\Delta}B)}}{\ctxext{(\ctxwk{\Gamma}{\Delta})}(\ctxwk{\Gamma}{B})}}
    {\typefont{wkext}_0(\Gamma,B)}}
\end{infarray}

\subsection{More stuff in the type theory}
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{A}}{\idfunc[A]}}
\end{infarray}

@@

Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$.
We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
$\ctxext{\Gamma}A$ and thus we have the inference
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\ctxwk{A}{P}}{\ctxwk{\ctxwk{A}{P}}{(\ctxwk{\ctxwk{A}{A}}{P})}}}
\end{equation*}
This family plays the role of the family $\jtype{\Gamma,x,y:A}{P(x)\to P(y)}$
in the type theory of \cite{TheBook}.

@@

\begin{defn}
Let $f$ be a term of $\ctxwk{A}{B}$ in context $\Gamma$.
\begin{enumerate}
\item For a family $Q$ over $B$ in context $\Gamma$ we define $f^\ast Q\jdeq\subst{f}{(\ctxwk{A}{Q})}$.
\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}B$ we define $f^\ast g\jdeq\subst{f}{(\ctxwk{A}{g})}$.
\end{enumerate} 
\end{defn}
\begin{rmk}
The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}A$ is can be viewed as the
type of functions from $A$ to $C$ which factor through $f$. Thus there should be
a function from $f^\ast(\ctxwk{B}{C})$ to $\ctxwk{A}{C}$.
\end{rmk}

\begin{rmk}
Every term $\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}B}{Q}}{\jtype{\ctxext{\Gamma}A}{f^\ast Q}}
\qquad
\inference{\jterm{\ctxext{\Gamma}B}{Q}{g}}{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{f^\ast g}}
\end{equation*}
\end{rmk}

\begin{defn}
A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$ is said to be a trivial
cofibration if we can infer
\begin{equation*}
\inference{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}B}{Q}{\tilde{t}}}\qquad
\inference{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}A}{f^\ast Q}{f^\ast \tilde{t}}{t}}
\end{equation*}
{\color{red}This statement should be reformulated so that it only involves a single judgment...
but I don't see directly how to do that.}
\end{defn}

@@

Another approach would be to define $f:\ctxwk{A}{B}$ to be left invertible
if there is a term $\ctxext({\Gamma}{A})P$

@@

\subsection{The basic constructors of type theory}
\subsubsection{Dependent function types}
\begin{infarray}{cc}
\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
& \inference{\jtypeeq{\ctxext{\Gamma}A}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
\inference{\jterm{\ctxext{\Gamma}A}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
& \inference{\jtermeq{\ctxext{\Gamma}A}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}A}{P}{\tfev(g)}}
& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}A}{P}{\tfev(g)}{\tfev(g')}}\\
\inference{\jterm{\ctxext{\Gamma}A}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
\end{infarray}

@@

\subsection{Identity types}
We need to define the diagonal function $\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{(\ctxext{A}\ctxwk{A}{A})}}{\delta}$
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{A}}{\idtypevar{A}}}\\
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}A}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
\inference{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}}{P}\qquad
\jterm{\ctxext{\Gamma}A}{\pairr{\delta,\reflsym}^\ast P}{d}}
{\jterm{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}}{P}{J}}
\end{infarray}

@@

the context $\tfext^\Gamma_0(A)$ of $\mfM$ by $\ctxext{\Gamma}A$.

@@

$\ctxext({\Gamma}{A})P$ for $\ctxext({\Gamma}{A})P$.

@@

\mftyp{\mftyp{\mfM}{\Gamma}}{A} & \jdeq\mftyp{\mfM}{\ctxext{\Gamma}A}\\
\mftyp{\tfext^\Gamma}{A} & \jdeq \modelfont{id}_{\mftyp{\mfM}{\ctxext{\Gamma}A}}\\

@@

\mftyp{\mftyp{\mfM}{\Gamma}}{A}\jdeq\mftyp{\mfM}{\ctxext{\Gamma}A},

@@

context in the model $\mftyp{\mfM}{\ctxext{\Gamma}A}$.

@@

\tfext^{\protect{\mftyp{\mfM}{\ctxext{\Gamma}A}}}(P,Q)

@@

\tfwk^A:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}

@@

\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}A}\to\mftyp{\mfM}{\Gamma}

@@

\item A context $\unit^\mfM:\tfctx(\mfM)$ and a term of type $\isequiv(\ctxext{\unit^\mfM}\blank)$. We will denote

@@

an identification $\alpha_\unit(\Gamma):\id{\ctxext{\Gamma}\unit^\Gamma}{\Gamma}$

@@

Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$.
We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
$\ctx{\Gamma}{A}$ and thus we can form the family
\begin{equation*}
\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}.
\end{equation*}
This family plays the role of the family $x,y:A\vdash P(x)\to P(y):\type$.
{\color{red}Explain this by performing the relevant substitution.}
Transportation is a term of the family
\begin{equation*}
\idtypevar{A}\to\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}
\end{equation*}
obtained by identity elimination, as usual. Thus we obtain a term
\begin{equation*}
\mathcal{T}rans_{x,y}:\terms{\id{x}{y}\to\subst{x}{P}\to\subst{y}{P}}.
\end{equation*}
for every $x,y:A$.

>>>>

\subsection{Extension}
We introduce extension which not only extends a context $\Gamma$ and a type
$A$ over it to a context $\ctxext{\Gamma}{A}$, but which also extends a type $A$
in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}{P}$ in context
$\Gamma$. We do this to ensure that all of type theory can be done in a context.
For instance, we could say (1) that a context in context $\Gamma$ is the same thing
as a type in context $\Gamma$; (2) When $A$ is a context in this sense, a type in
context $A$ is the same thing as a family $P$ over $A$ and (3) when $P$ is a type
in context $A$ in this sense, a term of $P$ keeps its original meaning.

Note that by introducing extension on the level of types and families,
we introduce $\Sigma$-types at a very early stage.
However, we need substitution to make this precise.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\ctxext{A}{P}}}
& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
\end{infarray}

\subsection{Weakening}
We first define weakening by a context $\Gamma$. Since weakening by $\Gamma$
should in principle be a functor, it acts on contexts, types and terms alike.
Note that it is because of the weakening $\ctxwk{\Gamma}{\Delta}$ of a context
$\Delta$ by $\Gamma$ that we can speak of context morphisms from $\Gamma$ to $\Delta$: they are the terms
of $\ctxwk{\Gamma}{\Delta}$.
\begin{infarray}{cc}
\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
\end{infarray}

A weakening operation is also defined for types. When $A$ and $B$ are both types
 in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$
 is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
 are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}{y}$ are the
 constant maps at $y$ from $A$ to $B$, for $y:B$.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}{B}}{Q}}
{\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}{B}}{Q}{Q'}}
{\jtypeeq{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
\end{infarray}

\subsection{Substitution}
\begin{infarray}{cc}
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}}{\jtype{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{{\Gamma}{A}}{P}}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}{\jterm{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{{\Gamma}{A}}{P}}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
\end{infarray}

@@

Extension is compatible with extension:
\begin{infarray}{c}
\inference
{ \jtype{\ctxext{\Gamma}{A}}{P}
}{\jterm
  {\ctxext{\Gamma}{{A}{P}}}
  {\ctxwk{\ctxext{\Gamma}{{A}{P}}}{(\ctxext{{\Gamma}{A}}{P})}}{\typefont{extext}_0(A,P)}
}\\
\inference
{ \jtype{\ctxext{{\Gamma}{A}}{P}}{Q}
}{\jterm
  {\ctxext{\Gamma}{{A}{{P}{Q}}}}
  {\ctxwk{\ctxext{A}{{P}{Q}}}{(\ctxext{{A}{P}}{Q})}}
  {\typefont{extext}_1(A,P,Q)}
}
\end{infarray}

Weakening is compatible with extension:
\begin{infarray}{c}
\inference{
  \jctx{\Gamma}
  \qquad
  \jtype{\Delta}{B}}
  {\jterm
    {\ctxext{\Gamma}{\ctxwk{\Gamma}{(\ctxext{\Delta}{B})}}}
    {\ctxwk{{\Gamma}{(\ctxext{\Delta}{B})}}{\ctxext{(\ctxwk{\Gamma}{\Delta})}{(\ctxwk{\Gamma}{B})}}}
    {\typefont{wkext}_0(\Gamma,B)}}
\end{infarray}

Substitution is compatible with substitution:
\begin{infarray}{c}
\inference{\jtype{\ctxext{{\Gamma}{A}}{P}}{Q}\qquad\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}{A}}{P}{f}}
{\jtypeeq{\Gamma}{\subst{x}{\subst{f}{Q}}}{\subst{{x}{f}}{\subst{x}{Q}}}}
\end{infarray}

\subsection{More stuff in the type theory}
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
\end{infarray}

@@

Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}{A}$.
We also have the family $\ctxwk{{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
$\ctxext{\Gamma}{A}$ and thus we have the inference
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\ctxwk{A}{P}}{\ctxwk{{A}{P}}{(\ctxwk{{A}{A}}{P})}}}
\end{equation*}
This family plays the role of the family $\jtype{\Gamma,x,y:A}{P(x)\to P(y)}$
in the type theory of \cite{TheBook}.

@@

\begin{defn}
Let $f$ be a term of $\ctxwk{A}{B}$ in context $\Gamma$.
\begin{enumerate}
\item For a family $Q$ over $B$ in context $\Gamma$ we define $f^\ast Q\jdeq\subst{f}{(\ctxwk{A}{Q})}$.
\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}{B}$ we define $f^\ast g\jdeq\subst{f}{(\ctxwk{A}{g})}$.
\end{enumerate} 
\end{defn}
\begin{rmk}
The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}{A}$ is can be viewed as the
type of functions from $A$ to $C$ which factor through $f$. Thus there should be
a function from $f^\ast(\ctxwk{B}{C})$ to $\ctxwk{A}{C}$.
\end{rmk}

\begin{rmk}
Every term $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{\Gamma}{A}}{f^\ast Q}}
\qquad
\inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
\end{equation*}
\end{rmk}

\begin{defn}
A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ is said to be a trivial
cofibration if we can infer
\begin{equation*}
\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
\end{equation*}
{\color{red}This statement should be reformulated so that it only involves a single judgment...
but I don't see directly how to do that.}
\end{defn}

@@

Another approach would be to define $f:\ctxwk{A}{B}$ to be left invertible
if there is a term $\ctxext{{\Gamma}{A}}{P}$

@@

\subsection{The basic constructors of type theory}
\subsubsection{Dependent function types}
\begin{infarray}{cc}
\inference{\jtype{\ctxext{\Gamma}{A}}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
& \inference{\jtypeeq{\ctxext{\Gamma}{A}}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
& \inference{\jtermeq{\ctxext{\Gamma}{A}}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}{A}}{P}{\tfev(g)}}
& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\tfev(g)}{\tfev(g')}}\\
\inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
\end{infarray}

@@

\subsection{Identity types}
We need to define the diagonal function $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{(\ctxext{A}{\ctxwk{A}{A}})}}{\delta}$
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}\\
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
\inference{\jtype{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}\qquad
\jterm{\ctxext{\Gamma}{A}}{\pairr{\delta,\reflsym}^\ast P}{d}}
{\jterm{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{A}}}{\idtypevar{A}}}{P}{J}}
\end{infarray}

@@

the context $\tfext^\Gamma_0(A)$ of $\mfM$ by $\ctxext{\Gamma}{A}$.

@@

$\ctxext{{\Gamma}{A}}{P}$ for $\ctxext{{\Gamma}{A}}{P}$.

@@

\mftyp{\mftyp{\mfM}{\Gamma}}{A} & \jdeq\mftyp{\mfM}{\ctxext{\Gamma}{A}}\\
\mftyp{\tfext^\Gamma}{A} & \jdeq \modelfont{id}_{\mftyp{\mfM}{\ctxext{\Gamma}{A}}}\\

@@

\mftyp{\mftyp{\mfM}{\Gamma}}{A}\jdeq\mftyp{\mfM}{\ctxext{\Gamma}{A}},

@@

context in the model $\mftyp{\mfM}{\ctxext{\Gamma}{A}}$.

@@

\tfext^{\protect{\mftyp{\mfM}{\ctxext{\Gamma}{A}}}}(P,Q)

@@

\tfwk^A:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}{A}}

@@

\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}{A}}\to\mftyp{\mfM}{\Gamma}

@@

\item A context $\unit^\mfM:\tfctx(\mfM)$ and a term of type $\isequiv(\ctxext{\unit^\mfM}{\blank})$. We will denote

@@

an identification $\alpha_\unit(\Gamma):\id{\ctxext{\Gamma}{\unit^\Gamma}}{\Gamma}$

@@

Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}{A}$.
We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
$\ctx{\Gamma}{A}$ and thus we can form the family
\begin{equation*}
\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}.
\end{equation*}
This family plays the role of the family $x,y:A\vdash P(x)\to P(y):\type$.
{\color{red}Explain this by performing the relevant substitution.}
Transportation is a term of the family
\begin{equation*}
\idtypevar{A}\to\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}
\end{equation*}
obtained by identity elimination, as usual. Thus we obtain a term
\begin{equation*}
\mathcal{T}rans_{x,y}:\terms{\id{x}{y}\to\subst{x}{P}\to\subst{y}{P}}.
\end{equation*}
for every $x,y:A$.
